---
title: Lin's Estimator
output: pdf_document
---
```{r}
library(haven)
library(tidyverse)
library(lmtest)
library(sandwich)

#Load dataset
AER <- read_dta("AER merged.dta")
```

Centering covariates and cleaning NAs:
```{r}
#Creating de-meaning function
demean <- function(x) {
  return(x-mean(x, na.rm=T))
}

#Covariates to demean
covs <- c("hpa", "freq", "years", "mrm2", "dormant",
          "female", "couple", "nonlit", "cases", "red0",
          "redcty", "pwhite", "pblack", "page18_39", "ave_hh_sz",
          "median_hhincome", "powner", "psch_atlstba", "pop_propurban")

#Apply de-meaning function
AER[covs] <- sapply(AER[covs], demean) 
AER_centered <- AER

#Remove NA values
AER_centered <- na.omit(AER_centered)
```

Function to calculate basic lin's estimator result
```{r}
#Input: data frame, treatment variable, outcome variable
#Output: estimator value, SE, p-value, R-squared
lin <- function(dat, treatment, outcome) {
  fit <- lm(paste(outcome, "~", treatment, "+hpa*", treatment, "+freq*", treatment,
                  "+years*", treatment, "+mrm2*", treatment, "+dormant*", treatment,
            "+female*", treatment, "+couple*", treatment, "+nonlit*", treatment,
            "+cases*", treatment, "+red0*", treatment, "+redcty*", treatment,
           "+pwhite*", treatment, "+pblack*", treatment, "+page18_39*", treatment,
            "+ave_hh_sz*", treatment, "+median_hhincome*", treatment,
            "+powner*", treatment, "+psch_atlstba*", treatment,
            "+pop_propurban*", treatment), data=dat)
return(c(fit$coefficients[2], sqrt(vcovHC(fit, type="HC1")[2,2]), coeftest(fit, vcovHC(fit, type="HC1"))[122], summary(fit)$r.squared))
}
```

```{r}
lin(AER_centered, "treatment", "gave")
lin(AER_centered, "treatment", "amount")
lin(filter(AER_centered, ratio!=2, ratio!=3), "ratio", "gave")
lin(filter(AER_centered, ratio!=2, ratio!=3), "ratio", "amount")
lin(filter(AER_centered, ratio!=1, ratio!=3), "ratio2", "gave")
lin(filter(AER_centered, ratio!=1, ratio!=3), "ratio2", "amount")
lin(filter(AER_centered, ratio!=1, ratio!=2), "ratio3", "gave")
lin(filter(AER_centered, ratio!=1, ratio!=2), "ratio3", "amount")
```

Blue or red state:
```{r}
lin(filter(AER_centered, red0<0), "treatment", "amount")
lin(filter(AER_centered, red0>0), "treatment", "amount")
lin(filter(AER_centered, red0<0), "treatment", "gave")
lin(filter(AER_centered, red0>0), "treatment", "gave")
```

Donated in 2005?
```{r}
lin(filter(AER_centered, dormant>0), "treatment", "gave")
lin(filter(AER_centered, dormant<0), "treatment", "gave")
lin(filter(AER_centered, dormant>0), "treatment", "amount")
lin(filter(AER_centered, dormant<0), "treatment", "amount")
```

Threshold
```{r}
lin(filter(AER_centered, size !=2, size !=3), "size25", "gave")
lin(filter(AER_centered, size !=2, size !=3), "size25", "amount")
lin(filter(AER_centered, size !=1, size !=3), "size50", "gave")
lin(filter(AER_centered, size !=1, size !=3), "size50", "amount")
lin(filter(AER_centered, size !=1, size !=2), "size100", "gave")
lin(filter(AER_centered, size !=1, size !=2), "size100", "amount")
```

Example amount
```{r}
lin(filter(AER_centered, ask !=2, ask !=3), "askd1", "gave")
lin(filter(AER_centered, ask !=2, ask !=3), "askd1", "amount")
lin(filter(AER_centered, ask !=1, ask !=3), "askd2", "gave")
lin(filter(AER_centered, ask !=1, ask !=3), "askd2", "amount")
lin(filter(AER_centered, ask !=1, ask !=2), "askd3", "gave")
lin(filter(AER_centered, ask !=1, ask !=2), "askd3", "amount")
```

Calculating treatment vs treatment results
```{r}
#3:1 vs 1:1
lin(filter(AER_centered, ratio!=2, ratio!=0), "ratio3", "gave")
lin(filter(AER_centered, ratio!=2, ratio!=0), "ratio3", "amount")

#2:1 vs 1:1
lin(filter(AER_centered, ratio!=0, ratio!=3), "ratio2", "gave")
lin(filter(AER_centered, ratio!=0, ratio!=3), "ratio2", "amount")

#3:1 vs 2:1
lin(filter(AER_centered, ratio!=1, ratio!=0), "ratio3", "gave")
lin(filter(AER_centered, ratio!=1, ratio!=0), "ratio3", "amount")
```
```{r}
#High threshold vs low threshold
lin(filter(AER_centered, size !=2, size !=0), "size100", "gave")
lin(filter(AER_centered, size !=2, size !=0), "size100", "amount")

#Medium threshold vs low threshold
lin(filter(AER_centered, size !=0, size !=3), "size50", "gave")
lin(filter(AER_centered, size !=0, size !=3), "size50", "amount")

#High threshold vs medium threshold
lin(filter(AER_centered, size !=0, size !=1), "size100", "gave")
lin(filter(AER_centered, size !=0, size !=1), "size100", "amount")
```


```{r}
#High ask vs low ask
lin(filter(AER_centered, ask !=2, ask !=0), "askd3", "gave")
lin(filter(AER_centered, ask !=2, ask !=0), "askd3", "amount")

#High ask vs medium ask
lin(filter(AER_centered, ask !=1, ask !=0), "askd3", "gave")
lin(filter(AER_centered, ask !=1, ask !=0), "askd3", "amount")

#Medium ask vs low ask
lin(filter(AER_centered, ask !=0, ask !=3), "askd2", "gave")
lin(filter(AER_centered, ask !=0, ask !=3), "askd2", "amount")
```

Below: attempts to use a generalized Lin's estimator employing various ML tools


1. Using a logistic regression method to predict `gave`:
```{r}
set.seed(2)
AER_centered_t <- filter(AER_centered, treatment==1)
AER_centered_c <- filter(AER_centered, treatment==0)
t_train <- sample(1:nrow(AER_centered_t), 0.8*nrow(AER_centered_t))
c_train <- sample(1:nrow(AER_centered_c), 0.8*nrow(AER_centered_c))
```

```{r}
library(MASS)

divide_by_sd <- function(x) {
  return(x/sd(x))
}
AER_cen_std <- AER_centered
AER_cen_std[covs] <- sapply(AER_centered[covs], divide_by_sd)
AER_cen_std_t <- filter(AER_cen_std, treatment==1)
AER_cen_std_c <- filter(AER_cen_std, treatment==0)
```

Fitting $\hat{\mu}_1$ model:
```{r}
glm.fit1=glm(gave~treatment+freq+years+mrm2+dormant+female+couple+nonlit+
              cases+red0+pwhite+page18_39+ave_hh_sz+
              powner+psch_atlstba+pop_propurban,
            data=AER_cen_std_t, subset=t_train, family=binomial)
glm_prob1 <- predict(glm.fit1, AER_cen_std_t[-t_train,], type="response")
glm_pred1 <- rep(0, 6204)

#Tuning probability threshold to maximize specificity and maintain accuracy > 90%
glm_pred1[glm_prob1 > 0.04]=1

#Testing model performance
confusionMatrix(as.factor(glm_pred1), as.factor(AER_cen_std_t[-t_train,]$gave))

#Full set predictions
glm_probs1 <- predict(glm.fit1, AER_cen_std_t, type="response")
glm_preds1 <- rep(0, 31018)
glm_preds1[glm_probs1 > 0.04]=1
```

Fitting $\hat{\mu}_0$ model:
```{r}
glm.fit0=glm(gave~treatment+freq+years+mrm2+dormant+female+couple+nonlit+
              cases+red0+pwhite+page18_39+ave_hh_sz+
              powner+psch_atlstba+pop_propurban,
            data=AER_cen_std_c, subset=c_train, family=binomial)
glm_prob0 <- predict(glm.fit0, AER_cen_std_c[-c_train,], type="response")
glm_pred0 <- rep(0, 3099)

#Tuning probability threshold to maximize specificity and maintain accuracy > 90%
glm_pred0[glm_prob0 > 0.031]=1

#Testing model performance
confusionMatrix(as.factor(glm_pred0), as.factor(AER_cen_std_c[-c_train,]$gave))

#Full set predictions
glm_probs0 <- predict(glm.fit0, AER_cen_std_c, type="response")
glm_preds0 <- rep(0, 15495)
glm_preds0[glm_probs0 > 0.031]=1
```

Calibrating predictions
```{r}
glm_preds0_cal <- glm_preds0+(1/nrow(AER_cen_std_c))*
                    sum(AER_cen_std_c$gave-glm_preds0)
glm_preds1_cal <- glm_preds1+(1/nrow(AER_cen_std_t))*
                    sum(AER_cen_std_t$gave-glm_preds1)

#Out-of-sample calibrated predictions
glm_probs0_oos <- predict(glm.fit0, AER_cen_std_t, type="response")
glm_preds0_oos <- rep(0, 31018)
glm_preds0_oos[glm_probs0_oos > 0.031]=1

glm_probs1_oos <- predict(glm.fit1, AER_cen_std_c, type="response")
glm_preds1_oos <- rep(0, 15495)
glm_preds1_oos[glm_probs1_oos > 0.04]=1

glm_preds0_oos_cal <- glm_preds0_oos+(1/nrow(AER_cen_std_c))*
                    sum(AER_cen_std_c$gave-glm_preds0)
glm_preds1_oos_cal <- glm_preds1_oos+(1/nrow(AER_cen_std_t))*
                    sum(AER_cen_std_t$gave-glm_preds1)
```

Resulting estimator
```{r}
(1/nrow(AER_centered))*(sum(AER_cen_std_t$gave-glm_preds0_oos)-
                          sum(AER_centered_c$amount-glm_preds1_oos))
```

2. Using xgboost to predict `gave`:
```{r}
library(xgboost)
vars <- c(covs, "treatment")
xg.fit <- xgboost(data=as.matrix(AER_centered_t[t_train,][vars]),
                  label=AER_centered_t[t_train,]$gave,
                  max.depth=4, nthread=2, nrounds=4,
                  objective="binary:logistic")
xg.pred <- predict(xg.fit, as.matrix(AER_centered_t[-t_train,][vars]))
xg.prediction <- as.factor(as.numeric(xg.pred > 0.1671))
confusionMatrix(xg.prediction, as.factor(AER_centered_t[-t_train,]$gave))

xg.fit2 <- xgboost(data=as.matrix(AER_centered_t[t_train,][vars]),
                  label=AER_centered_t[t_train,]$amount,
                  max.depth=4, nthread=2, nrounds=2)
xg.pred2 <- predict(xg.fit2, as.matrix(AER_centered_t[-t_train,][vars]))
plot(xg.pred2, AER_centered_t[-t_train,]$amount)
```

3. Using random forest to predict `gave`:
```{r}
set.seed(1)
library(randomForest)
rf <- randomForest(
             x=AER_centered_t[t_train,][covs],
             y=AER_centered_t[t_train,]$amount,
             xtest=AER_centered_t[-t_train,][covs],
             ytest=AER_centered_t[-t_train,]$amount, ntree=1000, mtry=19)
plot(rf$test$predicted, as.factor(AER_centered_t[-t_train,]$amount))
```

Using lasso to predict `gave`:
```{r}
library(glmnet)

#Mu_1
lasso_cv_1 <- cv.glmnet(model.matrix(amount ~ hpa + freq +
            years + mrm2 + dormant + female + couple + nonlit +
            cases + red0 + redcty + pwhite + pblack + page18_39 + 
            ave_hh_sz + median_hhincome + powner + psch_atlstba +
            pop_propurban, data=AER_centered_t[x_train,]),
                      AER_centered_t[x_train,]$amount, alpha=1)

lasso_preds_1 <- predict(lasso_cv, newx=model.matrix(amount ~ hpa + freq +
            years + mrm2 + dormant + female + couple + nonlit +
            cases + red0 + redcty + pwhite + pblack + page18_39 + 
            ave_hh_sz + median_hhincome + powner + psch_atlstba +
            pop_propurban, data=AER_centered_t[-x_train,]), s="lambda.min")

mean((lasso_preds_1-AER_centered_t[-x_train,]$amount)^2)

#Mu_0
AER_centered_c <- filter(AER_centered, treatment==0)
x_train_0 <- sample(1:nrow(AER_centered_c), 0.8*nrow(AER_centered_c))

lasso_cv_0 <- cv.glmnet(model.matrix(amount ~ hpa + freq +
            years + mrm2 + dormant + female + couple + nonlit +
            cases + red0 + redcty + pwhite + pblack + page18_39 + 
            ave_hh_sz + median_hhincome + powner + psch_atlstba +
            pop_propurban, data=AER_centered_c[x_train_0,]),
                      AER_centered_c[x_train_0,]$amount, alpha=0.5)

lasso_preds_0 <- predict(lasso_cv, newx=model.matrix(amount ~ hpa + freq +
            years + mrm2 + dormant + female + couple + nonlit +
            cases + red0 + redcty + pwhite + pblack + page18_39 + 
            ave_hh_sz + median_hhincome + powner + psch_atlstba +
            pop_propurban, data=AER_centered_c[-x_train_0,]), s="lambda.min")

mean((lasso_preds_0-AER_centered_c[-x_train_0,]$amount)^2)

mean_error_0 <-  mean(AER_centered_c[-x_train_0,]$amount-lasso_preds_0)
mean_error_1 <-  mean(AER_centered_t[-x_train,]$amount-lasso_preds_1)
```
